Inattentive responding can induce spurious associations between task behaviour and symptom measures

### 2021-02-23 manuscripts/main.tex
manuscript update because if i lose this i will never forgive myself

% load packages
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{xr-hyper} % for cross-referencing between documents
\usepackage{hyperref}
\usepackage{apacite}
\usepackage{float}
\usepackage{amsmath}
\usepackage[flushmargin]{footmisc}
\usepackage[flushleft]{threeparttable}
\usepackage{array}
\usepackage[font=small, labelfont=it]{caption}
\usepackage{booktabs} % for pretty tables

@@

% specify bibliography and link style
\bibliographystyle{apacite}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}
\renewcommand\bibliographytypesize{\small}

@@

\section{Main}

Paragraph 1: Rise of online studies. Utility for clinical research. Need to define TASK somewhere here.

Paragraph 2: Drawbacks to online studies. Noise. Data quality overview. Focus on unsystematic noise.

Paragraph 3: Noise is not systematic. As has been pointed out, careless responding can overinflate prevalence of symptoms and can yield spurious correlations. OR Need to define TASK somewhere here.

Paragraph 4: Not clear the extent to which this is an issue. For example, if screening in both types of data, might be not an issue. Even without screening, behavioral screening might be enough. Finally, behavior is complex and might not show up.

Paragraph 5: our current results call into question.

@@

\subsection{Screening for C/IE responding is common for task behavior but not for self-report behavior}

The presence of C/IE responding in a dataset can result in spurious correlations between measures of behavior. As such, it is imperative for researchers to detect and remove those participants whom complete online experiments with insufficient effort. The extent to which experimenters evaluate the quality of both the task and self-report data they collect is unclear. We suspect that whereas screening of task data is standard practice, the screening of self-report data is far less prevalent in online human behavioral research.

To provide quantitative support to this claim, we performed a cursory literature review of online human behavioral studies. (See Methods for details of the literature search and the list of studies included.) Briefly, we identified studies for inclusion through searches on Google Scholar using permutations of query terms related to online labor platforms, experimental paradigms, and symptom measures. We included studies that (a) recruited participants online, (b) measured behavior on at least one experimental task, and (c) measured behavior on at least one self-report symptom measure. This resulted in the inclusion of 49 studies. We then evaluated whether and how those studies performed task and self-report behavior screening. We note that this review was not meant to be systematic, but instead to provide a representative overview of the screening practices common to online behavioral studies.

Summary statistics about the prevalence and types of screening practices are provided in Table~\ref{tab:screening}. Of the 49 online behavioral studies meeting inclusion criteria, roughly 80\% (39/49) used at least one approach to identify C/IE responding in task behavior. Of those studies, just over half relied on a single evaluative method. As might be expected, there was considerable heterogeneity in task behavior screening methods. Most common was identifying participants whose performance was indistinguishable from chance-level on some measure of accuracy. Almost as common was screening based on response variability, or identifying participants who predominantly responded in the same fashion (e.g. used only 1 key). Less common approaches were screening based on response times and instruction comprehension questions.

In contrast, only the minority of online behavioral studies in our sample used at least one approach to identify C/IE responding in self-report behavior. Of those studies the most common method was the use of attention checks, which operate under the assumption that certain responses are unlikely under inattentive responding. Attention checks can be subdivided into different categories such as instructed items, where participants are explicitly told which response to select, and infrequency items, where only one or a small set of responses are valid. Of those studies that specified what type of attention check was used, instructed items were the most common form of attention check. Only a handful of studies employed statistical or so-called unobtrusive screening methods such as outlier detection or personal consistency.

In sum whereas screening for C/IE responding in task behavior is ostensibly the norm for online behavioral studies, screening self-report behavior is far less prevalent. One possible explanation for this pattern of results is that researchers (implicitly) assume that screening both task and self-report behavior is redundant. That is, researchers may believe that task and self-report screening methods will identify the same participants as having exhibited C/IE responding. To our knowledge, however, this crucial assumption has never been empirically verified. In the next section, we explicitly test this hypothesis in a large sample of online participants.

\begin{table}[t]
\centering
\setlength{\tabcolsep}{12pt}
\begin{tabular}{lrrrr}
\toprule
& \multicolumn{2}{c}{Task Screening} & \multicolumn{2}{c}{Self-Report Screening} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
Frequency & \multicolumn{2}{c}{N=39 (80\%)} & \multicolumn{2}{c}{N=19 (39\%)} \\
\midrule
Method & Accuracy & 18 (46\%) & Attention Check & 17 (89\%) \\
& Variability & 15 (38\%) & Instructed & 10 (53\%) \\
& Response Time & 7 (18\%) & Unspecified & 5 (26\%) \\
& Comprehension & 5 (13\%) & Unobtrusive & 4 (21\%) \\
& Other & 16 (41\%) \\
\bottomrule
\end{tabular}
\caption{The prevalence and types of task and self-report behavior screening practices in a representative sample (N=49) of online behavioral studies.}
\label{tab:screening}
\end{table}

Before continuing, we wish to briefly highlight two additional results from our literature review. First, we found that the majority of studies used only one single method to detect C/IE responding. This is worrisome from the perspective of signal detection theory. As has been previously reported, measures of C/IE responding have different sensitivities: some measures will identify only the worst participants, whereas others will flag many more respondents. As such, using only a single measure runs the risk of false positives (in the case of high sensitivity) and false negatives (in the case of low sensitivity). Similarly, different measures are sensitive to different \emph{types} of C/IE responding. Using only one measure may thus only identify a subset of C/IE respondents. As has been discussed elsewhere, the use of multiple measures is encouraged to more accurately and sensitively identify C/IE responding in a dataset.

Second, we found that the majority of studies that did screen self-report data relied on instructed items. This is separately worrisome as a growing body of research disputes their efficacy for online studies. There is empirical evidence that these items are less sensitive online elsewhere. A quick search online reveals that instructed items are frequently discussed in forums used by workers in online labor markets. Most concerning, there is empirical evidence that of a subset of participants that screen for these items and respond randomly where else. We encourage other methods and, as above, a multitude of methods.

\subsection{Task and self-report measures of C/IE responding}

To measure the correspondence between measures of C/IE responding based on task and self-report behaviors, we conducted an behavioral experiment in two prominent online labor marketplaces (see Methods for complete details). In total, 386 total participants from Amazon Mechanical Turk (N=186) and Prolific (N=200) completed an online study comprised of one behavioral task and several self-report symptom measures. We recruited participants from both platforms in order to improve the generalizability of our findings. The demographics of each sample are reported in the Supplement.

Participants completed a 3-arm probabilistic reversal learning bandit task. During the task, participants were presented with three options and could only choose one, after which they received feedback (reward or no-reward). Participants were instructed to choose the option that yielded reward most often, which they would learn through trial-and-error. Feedback was probabilistic such that the best option on a given trial returned reward on 80\% of trials and the other options only 20\%. Importantly, the best option changed every 15 trials. Participants were made explicitly aware that the best option would change over time, but they were not told when or how often this would occur. Participants completed 90 trials of the task (5 total reversals).

Prior to the start of the behavioral experiment, participants also completed four standard self-report symptom measures: the generalized anxiety disorder scale (GAD7), the 7-up/7-down, the BIS/BAS, and the Snaith-Hamilton Pleasure Scale. The measures were selected to provide a range of response distributions. For example, the GAD-7 and 7-up/7-down are substantially asymmetric.

To measure the level of correspondence between task- and self-report-based measures of C/IE responding, we computed a number of metrics from the task and self-report behavior of each subject. For the task behavior, we estimated the following metrics: response variability, accuracy, win-stay lose-shift, and fraction of trials with suspicious response times. For the self-report behavior, we estimated the following metrics: number of suspicious responses to infrequency items, inter-item standard deviation (ISD), personal reliability, Mahalanobis distance, and the per-item reading time. Each metric was selected because it has previously been used.

To provide an overall sense of the quality of the data, we first focus on the task accuracy and the infrequency item metrics as these represent the most common screening methods of each behavioral modality. Only 26 participants (7\%) were flagged as exhibiting choice behavior at or above chance levels. In contrast, 85 participants (22\%) responded suspiciously to one or more of the infrequency items while completing of self-report symptom inventories. This discrepancy in the proportion of participants flagged by each method is consistent with previous studies which have found differing sensitivities to C/IE responding. Notably, the proportions of flagged participants were marginally but significantly greater on Mechanical Turk compared to Prolific for both accuracy ($z=2.224, p=0.026$) and infrequency ($z = 2.223, p = 0.026$) metrics.

Core result 1: continuous (spearman) correlaiton analysis.

\subsection{}

>>>>

% load packages
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{xr-hyper} % for cross-referencing between documents
\usepackage{hyperref}
\usepackage{float}
\usepackage{amsmath}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[flushmargin]{footmisc}
\usepackage[flushleft]{threeparttable}
\usepackage{array}
\usepackage[font=small, labelfont=it]{caption}
\usepackage{booktabs} % for pretty tables

@@

% specify link style
\addbibresource{main.bib}
\renewcommand*{\bibfont}{\small}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}

@@

\section{Introduction}

In recent years, online labour markets (e.g. Amazon Mechanical Turk, Prolific, Crowdflower) have become increasingly popular as a source of research participants in the behavioral sciences \cite{stewart2017crowdsourcing}, in no small part due to the ease with which these services allow for recruitment of large, diverse samples. The advantages of online data collection have also begun to be recognized in psychiatric research \cite{chandler2016conducting}, where this method offers several distinct advantages over traditional approaches to participant recruitment. The ability to assess psychiatric symptom severity in large general-population samples makes possible large-scale transdiagnostic analysis \cite{gillan2016taking, rutledge2019machine}, and facilitates recruitment from difficult-to-reach participant populations \cite{strickland2019use}. Online labour markets also facilitate re-recruitment, making them an attractive option for validating the psychometric properties of assessment tools \cite{enkavi2019large} or studying clinical processes longitudinally \cite{kothe2019retention}.

With the advantages of online data collection also come specific drawbacks. Since participants recruited from online labour markets are typically completing experiments in their homes, they may be more likely to be distracted or multi-tasking during an experiment. Similarly, unsupervised participants may be more likely to use careless/insufficient effort (C/IE) response strategies with the intention to minimise expenditure of time and cognitive effort (e.g., responding randomly on self-report surveys or behavioural tasks). Among researchers using online labour markets, a common view is that poor-quality data resulting from C/IE responding can simply be treated as a source of unsystematic measurement error that can be overcome with increased sample sizes \cite{gillan2016taking, chandler2020participant}. Similarly, it is common practice in online behavioural research to mitigate poor-quality data using the same screening methods that are typically used in in-person data collection (e.g., excluding participants who perform at- or below-chance on behavioural tasks).

Here we wish to draw attention to an underappreciated feature of online psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom distributions tend to be heavily positively skewed. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey accurately \cite{chandler2020participant, ophir2020turker}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient effort.

When participants complete both symptom surveys and behavioral tasks (a common study design in computational psychiatry), this artefact has the potential to induce spurious correlations between symptom self-reports and task behavior. Concretely, if the same participants who engage in C/IE responding on surveys (and who therefore inaccurately report high levels of psychiatric symptoms) also respond carelessly on behavioral tasks, then we would expect to observe an entirely spurious correlation emerge between greater symptom severity and worse task performance (see Figure \ref{fig:simulation}). A similar effect has been well documented in personality psychology, where the presence of careless respondents can induce correlations between questionnaires and bias estimated factors in factor analysis \cite{huang2012detecting, robinson2014inaccurate, huang2015insufficient, chandler2020participant, arias2020little}. To our knowledge, however, the potential for C/IE responding to induce spurious correlations between symptom endorsement and task behavior has not previously been noted in the burgeoning field of computational psychiatry.

\begin{figure}[t]
\includegraphics[width=16cm]{../figures/main_01.png}
\centering
\caption{Caption TK here}
\label{fig:simulation}
\end{figure}

Here we demonstrate the real risk that C/IE responding can lead to spurious symptom-task correlations in computational psychiatry research. First, we asked to what extent researchers in computational psychiatry screen self-report data from online samples. We found that the majority of these studies did not screen participants' survey data at all, and that very few followed best-practice recommendations for survey data screening. We then asked whether behavioural screening alone was sufficient to identify participants engaging in C/IE responding on psychiatric symptom surveys. In a new dataset from two separate online labour marketplaces, we found that screening based on task behaviour fails to adequately identify participants engaging in C/IE responding on surveys. Lastly, we investigated whether, under these circumstances, C/IE responding led to spurious correlations between symptom severity and task performance for positively-skewed symptom measures. Consistent with the logic set out above, we confirmed that failure to appropriately screen out C/IE survey responding in the proof-of-concept dataset that we collected would have produced a number of spurious correlations between task behaviour and self-reported symptoms of depression and anxiety.

@@

\subsection{Screening for C/IE responding is common for task behavior but not for self-report surveys}

When the rate of symptom endorsement is low in the general population, participants who engage in C/IE responding on self-report symptom measures may appear as putative symptomatic individuals. If the same participants also perform a behavioural task with insufficient effort, they may in turn induce spurious correlations between measures of task behavior and self-report symptoms. The risk for erroneous inferences, however, is attenuated if experimenters identify and remove those participants. As such, it is imperative for researchers to screen for C/IE responding in their data. The extent to which experimenters evaluate the quality of both the task and self-report data they collect is unclear. We suspect that whereas screening of task data is standard practice, the screening of self-report data is far less prevalent in online human behavioral research.

To test the empirical support for this claim, we performed a narrative literature review of online human behavioral studies (see Methods for details of the literature search and the Github repository for the full list of included studies). Briefly, we identified studies for inclusion through searches on Google Scholar using permutations of query terms related to online labour platforms, experimental paradigms, and symptom measures. We included studies that (a) recruited participants online, (b) measured behavior on at least one experimental task, and (c) measured responses on at least one self-report symptom measure. We then evaluated whether and how each of the resulting 49 studies performed task and self-report data screening. We note that this review was not meant to be systematic, but instead to provide a representative overview of common screening methods in online behavioral studies.

Summary statistics concerning the prevalence and types of screening methods are provided in Table~\ref{tab:screening}. Of the 49 online behavioral studies meeting inclusion criteria, roughly 80\% (39/49) used at least one method to identify C/IE responding in task behavior. Of those studies, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

In contrast, only a minority (19/49, or 39\%) of online behavioral studies in our sample screened for C/IE responding in self-report symptom measures. Of those studies the most common method was the use of attention checks, which operate under the assumption that certain responses to these prompts are unlikely under attentive responding, such that participants who do not give the correct response are likely to be engaged in C/IE responding. Attention checks can be subdivided into instructed items (in which participants are explicitly told which response to select; e.g., `Please select ``Strongly Agree"'), and infrequency items (in which some responses are logically invalid; e.g., endorsing 'Agree' for the question `I competed in the 1917 Summer Olympic Games'). Of those studies that specified what type of attention check was used, instructed items were the most common method. As we discuss further below, this is notable because best-practice recommendations for data collection in personality psychology explicitly counsel against using instructed-item attention checks. Only a handful of studies employed statistical or so-called unobtrusive screening methods such as outlier detection or personal consistency.

In sum, screening methods for C/IE responding vary widely across the studies that we reviewed. Whereas screening for C/IE responding in task behavior is relatively common for online behavioral studies, screening of self-report survey data is far less prevalent. Though at first this pattern may seem troubling, low rates of survey data screening are not necessarily an issue if screening on task behavior alone is sufficient to remove participants engaging in C/IE responding. That is, screening on survey data may be redundant if there is a high degree of correspondence between task- and survey-based screening methods. To our knowledge, however, this assumption has never been empirically tested. In the next section, we explicitly test this hypothesis in a large sample of online participants completing a battery of self-report surveys and a behavioural task.

\begin{table}[t]
\centering
\setlength{\tabcolsep}{8pt}
\begin{tabular}{lllllll}
\toprule
& \multicolumn{2}{c}{Task Screening} & \multicolumn{4}{c}{Self-Report Screening} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-7}
Frequency & \multicolumn{2}{c}{N=39 (80\%)} & \multicolumn{4}{c}{N=19 (39\%)} \\
\midrule
Measure & Accuracy & 18 (37\%) & \multicolumn{2}{l}{Attention Check} & & 17 (35\%) \\
& Variability & 15 (31\%) & & \textit{Instructed} & & 10 (20\%) \\
& Response Time & \ 7 (14\%) & & \textit{Unspecified} & & 5 (10\%) \\
& Comprehension Check & \ \ 5 (10\%) & \multicolumn{2}{l}{Unobtrusive} & & \ 4 (8\%) \\
& Other & 16 (33\%) \\
\bottomrule
\end{tabular}
\caption{The prevalence and types of task and self-report data screening practices in a sample (N=49) of online behavioral studies.}
\label{tab:screening}
\end{table}

Before continuing, we wish to briefly highlight two additional results from our literature review. First, we found that the majority of studies used only one single method to detect C/IE responding. This is worrisome from the perspective of signal detection theory. As has been previously reported \cite{desimone2018dirty}, measures of C/IE responding vary in their sensitivities: some measures will identify only the most careless participants, whereas others will flag many more participants. As such, using only a single measure runs the risk of false positives (in the case of high sensitivity) and false negatives (in the case of low sensitivity). Similarly, screening methods vary in their sensitivity to different patterns of C/IE responding \cite{curran2016methods}. Using only one measure may thus only identify a subset of careless respondents. As has been discussed elsewhere, the use of multiple screening methods is encouraged to more accurately and sensitively identify C/IE responding in a dataset \cite{curran2016methods, barends2019noncompliant}.

Second, we found that the majority of studies that did screen self-report survey data relied on instructed attention checks. This is potentially problematic, since a growing body of research disputes the efficacy of instructed items as attention checks for online studies \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Indeed, instructed attention checks are often discussed on forums for workers of online labour markets\footnote{Several examples of such discussions can be found at the Github repository for this project:  \url{github.com/nivlab/silver-screen}}, and scripts have even been developed with the purpose to highlight words most commonly used in instructed attention checks. Unsurprisingly, then, it has been documented that subsets of participants may actively seek out instructed items in surveys and respond randomly elsewhere \cite{barends2019noncompliant}. In line with best-practice recommendations for online survey research, therefore, we encourage experimenters to use alternate means of screening for C/IE responding for self-report survey data (e.g., infrequency-item attention checks) \cite{huang2015detecting}.

\subsection{Careless participants appear symptomatic when the overall level of symptom endorsement is low}

In the majority of the papers reviewed above, experimenters relied solely on behavioural screening to exclude participants who engaged in C/IE responding during an experiment. Nevertheless, even with behavioural screening, it is still possible that spurious behaviour-symptom correlations might emerge if there is low correspondence between behavioural and survey-based C/IE screening methods (since high correspondence between methods would entail that behavioural screening alone would exclude participants who would have failed survey screening).

In order to the measure the correspondence of screening measures estimated from task and self-report behavior, we conducted an online behavioral experiment involving a simple behavioral task and a battery of self-report psychiatric symptom measures (see Methods for details). A final sample of 386 participants from the Amazon Mechanical Turk (N=186) and Prolific (N=200) online labour markets completed a probabilistic reversal learning task and 5 self-report symptom measures. The reversal learning task required participants to learn through trial-and-error which of three choice options yielded reward most often, and was modeled after similar tasks used to probe reinforcement learning deficits in psychiatric disorders \cite{huang2017computational, waltz2007probabilistic, mukherjee_reward_2020}. The self-report measures assayed psychiatric symptoms related to depression, mania, anxiety, reward motivation, and anhedonia, and were chosen based on previous literature to have a variety of expected response distributions (symmetric and asymmetric). In line with current best-practice recommendations in personality psychology, each self-report instrument included one `infrequency' item that could be used to identify C/IE responses in survey data (see Methods for a list of infrequency items). To minimize any influence of fatigue, the entire experiment (surveys \& task) was designed to require 10 minutes on average to complete (observed mean = 10.28 minutes). For similar reasons, participants completed the surveys prior to beginning the task.

To assess the overall quality of the data, we examined the number of participants flagged by the choice accuracy and infrequency item screening measures (i.e. two most common screening measures identified in our literature review). Only 26 participants (7\%) were flagged as exhibiting choice behavior in the reversal learning task at or below chance levels. In contrast, 85 participants (22\%) responded suspiciously on one or more of the infrequency items when completing the self-report symptom measures. This discrepancy in the proportion of participants flagged by each method is consistent with previous research which have found varying levels of sensitivity to C/IE responding across screening methods \cite{desimone2018dirty}. Though we will not address platform differences here, we note the proportions of flagged participants were marginally but significantly greater on Mechanical Turk compared to Prolific for both accuracy (MTurk: N=18/186; Prolific: N=8/200; $z=2.22, p=0.026$) and infrequency (MTurk: 50/186; Prolific: 35/200; $z = 2.22, p = 0.026$) measures.

\begin{figure}[t!]
\includegraphics[width=16cm]{../figures/main_02a.png}
\centering
\caption{Caption TK here}
\label{fig:distributions}
\end{figure}

According to our hypothesis, a prerequisite for spurious behavior-symptom correlations to emerge is a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants, which might occur when the overall rate of symptom endorsement is low. As such, we visually inspected for each of the self-report measures the distribution of symptom scores in participants flagged by the infrequency measure vs. those that were not (hereafter referred to as `careless' or `attentive`, respectively; Figure \ref{fig:distributions}). In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in careless participants relative to attentive participants for the symptom measures where symptom scores were positively-skewed (e.g. 7-up, 7-down, GAD-7). In contrast, where there was a greater range of symptom endorsement, the distributions of symptom scores between the two groups of participants were less noticeably distinct.

\begin{table}[!t]
\centering
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lccccccc}
\toprule
& & \multicolumn{2}{c}{Total Score} & \multicolumn{2}{c}{Clinical Levels} & \multicolumn{2}{c}{Cronbach's $\alpha$} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6} \cmidrule(lr){7-8}
 Subscale & Skew & Attentive & Careless & Attentive & Careless & Attentive & Careless \\
\midrule
\midrule
7u    &  0.81 &    3.87 &   10.15 &     0.04 &     0.45 &    0.84 &    0.84 \\
7d    &  0.76 &    4.75 &   10.68 &     0.09 &     0.46 &    0.94 &    0.88 \\
GAD7  &  0.75 &    4.86 &    9.73 &     0.17 &     0.56 &    0.92 &    0.87 \\
BIS   &  0.78 &    7.74 &    7.92 &        - &        - &    0.83 &    0.62 \\
BAS   &  0.17 &   15.73 &   16.16 &        - &        - &    0.84 &    0.71 \\
SHAPS &  0.26 &    8.02 &   10.85 &     0.15 &     0.29 &    0.90 &    0.81 \\
PSWQ  &  0.19 &    4.78 &    6.74 &     0.07 &     0.08 &    0.93 &    0.81 \\
\bottomrule
\end{tabular}
\caption{}
\label{tab:screening}
\end{table}

To verify the infrequency item method as an accurate assay of C/IE responding, we performed three validation analyses. The first approach is to compare the proportion of participants meeting the cutoff for clinical psychopathology among careless versus attentive participants(Table \ref{tab:screening}). On the most positively-skewed measures, the fraction of participants scoring at or above clinical levels of symptom endorsement was dramatically inflated in careless responders. For example, previous studies of prevalence of generalized anxiety disorder found that roughly 5\% of the general population met the clinical threshold at a score of 10 or higher on the GAD-7 \cite{lowe2008validation, hinz2017psychometric}. Though clinical thresholds have not been established for the 7-up/7-down, recent work has suggested scores of 12 or higher as useful for discriminating between those with and without clinical levels of pathology \cite{youngstrom2020evaluating}. Though the fraction of participants meeting threshold for these measures in the attentive group was within the expected range of prevalence given epidemiological data, the exaggerated rates in the flagged group are suggestive of C/IE responding.\footnote{Interestingly, compared to previous literature the proportion of participants meeting threshold on the GAD-7 was elevated. We suspect this may reflect elevated rates of state anxiety during the COVID-19 pandemic, when this data was collected.}

The second approach was to compare the estimated internal consistency of the self-report measures between the careless and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the careless group (as has been previously reported; \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}). Accordingly, on the majority of measures we observed a reduction in Cronbach's $\alpha$ in the careless group relative to the attentive group (Table \ref{tab:screening}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance even controlling for the difference in the number of participants between the two groups ($t = 4.25, p = 0.006$).

The third and final approach is to statistically quantify the degree to which participants responded to the items in the self-report symptom inventory in a stereotyped fashion; that is, to determine if participants exhibited patterns in their responses that were independent of the contents of the items. To do so, we fit a random intercept item factor analysis model \cite{maydeu2006random} to the self report data (see Methods for details). Briefly, an intercept term is estimated for each participant which quantifies their bias towards using particular response options (e.g. leftmost, rightmost) regardless of what that option signifies for a particular self-report measure. We observed a credible difference between the average value of this intercept between the two groups (stats here), such that participants suspected of C/IE responding were biased towards using the right-half of the response options on the self-report symptom inventory. This translates to a tendency to endorse greater illness on the 7-up/7-down and GAD-7 (where the rightmost options indicate greater frequency of symptoms) but translates to a tendency to endorse \emph{less} illness on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition). This pattern also helps explain why the distributions of symptom scores between attentive and careless participants on the SHAPS appear to be similar despite the low rates of symptom endorsement usually observed on the SHAPS \cite{snaith1995scale}.

In sum, we found that almost a quarter of participants in our online sample were suspected of C/IE responding when screened using the infrequency-item measure, in contrast to only 7\% of participants flagged when using task accuracy as a screening measure. Importantly, participants suspected of C/IE responding were more likely to appear as symptomatic on self-report measures where symptom endorsement was rare on average. Furthermore, using three independent analyses, we identified evidence that validates the infrequency-item measures as sensitive to C/IE responding. These validation analyses also identified multiple forms of C/IE responding. Though the Cronbach's alpha result suggests at least some participants were engaging in random responding, the item response model analysis identified participants who exhibited a preference for the rightmost response options regardless of their semantic content. That C/IE responding in this data manifested in more than one way (consistent with previous reports \cite{curran2016methods}) further highlights the importance of utilizing multiple screening methods, a point we will return to in the discussion.

\subsection{Low correspondence between task and self-report measures of C/IE responding}

Next, we evaluated the degree of correspondence between behavioral and self-report screening measures. This allowed us to determine whether screening on behavior alone was sufficient to identify and remove careless participants. In line with the narrative review, we computed multiple measures of C/IE responding from each participant's task behaviour and survey responses (see Methods for full description). For task behaviour, we computed the following screening measures: response variability (the fraction of responses using only one key), accuracy (the fraction of choices of the reward-maximizing response option), win-stay lose-shift rates (the difference in likelihood of repeating the previous choice following rewarded vs. unrewarded trials), and fraction of excessively fast response times. In line with previous studies, for self-report responses we screened participants' data for C/IE responding using the following measures: infrequency items (the number of suspicious responses to infrequency items), inter-item standard deviation (ISD; the variability of responses around a participant's modal response), personal reliability (the correlation between split-halves of a participant's responses), Mahalonobis distance (the extent to which a participant is an outlier with respect to all participants), and reading time (the average time taken to respond to each item).

To measure the degree of correspondence between these behavioral and self-report screening measures, we performed two complementary analyses. First, we computed pairwise correlations on the unthresholded (continuous) measures using Spearman's rank correlation. The resulting pairwise similarity matrices are presented in Figure \ref{fig:correspondence} (left panel). After correcting for multiple comparisons, there were few significant correlations between the behavioral and self-report measures. Only choice accuracy showed significant associations with any self-report measure (specifically the infrequency and Mahalonobis distance measures). Crucially, the size of these observed correlations were roughly half that observed for the correlations between the self-report measures. This is worrisome as it suggests that, though there is some relation between C/IE responding on tasks and self-report inventories, it is not strong enough to reliably detect careless participants in self-report data from task data alone.

\begin{figure}[!t]
\includegraphics[width=16cm, trim=0 0 3cm 0]{../figures/main_02b.png}
\centering
\caption{Caption TK here}
\label{fig:correspondence}
\end{figure}

To complement the correlational analysis, we also estimated the pairwise level of agreement on the binarized measures using the Dice similarity coefficient. This second approach quantifies the degree of overlap between the set of would-be excluded participants based on different screening measures under a common exclusion rate. Though measures like accuracy have relatively clear threshold cutoffs (e.g. chance level performance), the majority of the measures evaluated here do not. As such, we evaluated the measures with respect to the top 10\% of suspicious participants flagged by each measure, corresponding roughly to the fraction of participants having performed at chance levels on the reversal learning task. The results of the same analysis repeated for the top 25\% of suspicious participants (corresponding roughly to the fraction of participants flagged by the infrequency item measure) produced similar results, and is included in the supplement.)

The resulting pairwise similarity matrices for each approach are presented in Figure \ref{fig:correspondence} (right panel). The pattern of results are largely consistent with the correlational analysis: few pairs of task and self-report screening measures achieved levels of agreement greater than what would be expected by chance. The only significant cross-modality pair identified -- between the infrequency item and choice accuracy measures -- has a similarly coefficient less than 0.4. In other words, when these two measures are used to identify the top 10\% most suspicious participants, they agree on only two out of every five participants. Screening on choice accuracy alone (the most common method identified in our narrative review) would fail to identify the majority of participants most likely engaging in C/IE responding as determined by the infrequency items.

To conclude this section, we found across two complementary analyses that measures of C/IE responding in task and self-report data were not redundant. That is, contrary to common implicit assumptions in online behavioural research, detecting and removing careless participants based on behavioral data alone would not be sufficient for ensuring quality self-report data in our sample. Though we cannot extrapolate from this one dataset to conclude that this will always be the case, we can nevertheless conclude that experimenters cannot simply assume that screening on self-report data is redundant with screening on task behavior (and vice-versa). While tendencies towards C/IE responding in tasks and self-report inventories may be associated, our results show that both types of screening are likely to be necessary to ensure acceptable data quality.

\subsection{Spurious symptom-behaviour correlations produced by C/IE responding}

Here we detail the potential consequences of screening only on task behavior in our data. To do this, we estimated the pairwise correlations between the symptom scores of each of the self-report measures and several measures of performance on the reversal learning task. This analysis emulated a typical computational psychiatry analysis, in which the results of primary interest are the correlations between task behaviour and self-reported psychiatric symptom severity.

For each participant, we computed both descriptive and model-based measures of behaviour on the reversal learning task. Descriptive measures included accuracy (the fraction of choices of the reward-maximizing response option), points (the total number of points accumulated over the game), win-stay rates (the fraction of trials on which a participant repeated the previous trial's choice following a reward outcome), lose-shift rates (the fraction of trials on which a participant deviated from the previous trial's choice following a non-reward outcome), perseveration (the number of trials on which a participant continued to choose the previously dominant response option following a reversal). The model-based measures included the parameters of commonly-used reinforcement model of choice behavior, i.e. $Q$-learning with asymmetric learning rates (see Method for further details): choice sensitivity ($\beta$,  inverse temperature), positive learning rate ($\eta_p$, learning rate following a positive reward prediction error), negative learning rate ($\eta_n$, learning rate following a negative reward prediction error), and learning rate asymmetry ($\kappa$, difference between $\eta_p$ and $\eta_n$ in a transformed space). We chose these task performance measures as they have been previously used to assess performance in clinical samples \cite{huang2017computational, brolsma2020challenging, mukherjee_reward_2020, ritschel2017neural}.

To understand the effects of applying different forms of screening, we estimated the correlations between each unique pairing of a self-report symptom measure and measure of behavior under four different conditions: no screening, screening only on task behavior (i.e. accuracy above chance), screening only on self-report responses (i.e. pass responses on all infrequency items), or both. The resulting pairwise behavior-symptom correlations following each screening procedure are presented in Figure \ref{fig:correlations}. To note, these correlational analyses were not corrected for multiple comparisons as the intended purpose here is to emulate a conventional analysis where fewer statistical tests would be performed, as dictated by one's hypotheses.

\begin{figure}[!t]
\includegraphics[width=16cm]{../figures/main_03a.png}
\centering
\caption{Caption TK here}
\label{fig:correlations}
\end{figure}

When no rejections based on for C/IE responding was applied (i.e. all participants in this sample were included; top left, Figure \ref{fig:correlations}), many significant correlations emerged between measures of task behavior and symptom scores. Consistent with our predictions, the majority of these correlations involved symptom measures with asymmetric total score distributions. Attending to only the most skewed measures (i.e. 7-up/7-down, GAD-7), symptom endorsement was correlated with almost every behavioral measure. That is, significant correlations were not restricted only to behavioral measures often used as proxies for participant effort (e.g. accuracy, inverse temperature $\beta$) but also to measures of previous theoretical interest. Conversely, we found few significant correlations among symptom measures with more symmetric distributions. This is despite the fact these scales measure similar symptoms and syndromes (e.g. anxiety \& worry, depression \& anhedonia).

Next, we rejected participants based on task behavior screening (i.e. choice accuracy, removing the 7\% of participants exhibiting behavior indistinguishable from chance; top right, Figure \ref{fig:correlations}). The pattern of correlations was largely unchanged: we again found many significant correlations between measures of behavior and asymmetric symptom measures, but almost no significant correlations involving symmetric symptom measures. Looking again at the most skewed symptom measures, all correlations involving accuracy and the inverse temperature were preserved. In sum, rejections based on the most common form of behavioral screening (i.e. performance accuracy) had little effect on behavior-symptom correlations as compared to no screening.
